Notes on Design and Implementation of Anglican

### 2015-07-29 paper/paper.tex
managing primitive procedures

>>>>

When an Anglican function is transformed into a Clojure function
by \texttt{fn-cps}, two auxiliary parameters are added to the
beginning of the parameter list --- continuation and state.
Correspondingly, when a function \textit{call} is transformed
(by \texttt{cps-of-application} or \texttt{cps-of-apply}), the
current continuation and the state are passed to the called
function. Anglican can also call Clojure functions; however
Clojure functions do not expect these auxiliary parameters.
To allow the mixing of Anglican (CPS-transformed) and Clojure
function calls in Anglican code, the Anglican compiler must be
able to recognize `primitive' (that is, implemented in Clojure
rather than in Anglican) functions.

Providing an explicit syntax for differentiating between
Anglican and Clojure function calls would be cumbersome. Another
option would be to use meta-data to identify Anglican function
calls at runtime, however this would impact performance, and
a good runtime performance is critical for probabilistic
programs. The approach taken by Anglican is to maintain a list
of unqualified names of primitive functions, as well of
namespaces in which all functions are primitive, and recognize
primitive functions by name --- if a function name is not in the
list, the function is a Clojure function. Global dynamically
bound variables \texttt{*primitive-procedures*} and
\texttt{*primitive-namespace*} contain the initial lists of
names and namespaces, correspondingly. Of course, local bindings
can shade global primitive function names. For example,
\texttt{first} is a Clojure function inside the let block in the
following example:
\begin{lstlisting}[frame=single,language=Anglican, basicstyle=\scriptsize\ttfamily, showstringspaces=false]
(let [first (fn [[x & y]] x)]
  (first '[1 2 3]))
\end{lstlisting}
The Anglican compiler takes care of the shading by rebinding
\texttt{*primitive-procedures*} in every lexical scope
(\texttt{fn-cps}, \texttt{cps-of-let}). Macro
\texttt{shading-primitive-procedures} automates the shading.
